### 2015-04-10 contract-language.tex
edits

The syntax of our core contract language is shown in Figure
~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
admits prenex universal quantification over typed and untyped effect variables.
We use a special effect variable ($\cureff$) to denote the effect of
\emph{current operation} - the operation for which a contract is being written.
Notice that $\cureff$ occurs free in the contract. We will fix its scope when
classifying the contracts (\S~\ref{sec:classify}). The type of an effect is
simply the name of the operation (eg: \rcf{withdraw}) that induced the effect.
We admit disjunction in types to let an effect variable range over multiple
operation names. The contract $\small \forall (a : \tau_1 \vee \tau_2).~\psi$
is just syntactic sugar for $\small \forall a. (\oper{a}{\tau_1} \vee
\oper{a}{\tau_2}) \Rightarrow \psi$. An untyped effect variable ranges over all
operation names.

Quantifier-free propositions in our contract language are conjunctions,
disjunctions and implications of predicates expressing relations between pairs
of effect variables. The syntactic class of relations is seeded with primitive
$\visZ$, $\soZ$, and $\sameobjZ$ relations, and also admits derived relations
that are expressible as union, intersection, or transitive
closure\footnote{Strictly speaking, $R^{+}$ is not the transitive closure of
$R$, as transitive closure is not expressible in FOL.  Instead, $R^{+}$ in our
language denotes \emph{a} superset of transitive closure of $R$. Formally,
$R^{+}$ is any relation $R'$ such that forall $x$, $y$, and $z$, a) $R(x,y)
\Rightarrow R'(x,y)$, and b) $R'(x,y) \conj R'(y,z) \Rightarrow R'(x,z)$} of
primitive relations.  Commonly used derived relations are the \emph{same object
session order} ($\small \sooZ = \soZ ~\cap~ \sameobjZ$), \emph{happens-before
order} ($\small \hbZ = (\soZ ~\cup~ \visZ)^+$) and the \emph{same object
happens-before order} ($\small \hboZ = (\sooZ ~\cup~ \visZ)^+$).

@@

\item \textbf{Causal consistency}: Causally consistent operations are required
	to see a causally consistent snapshot of the object state, including the
	actions performed on the same session.  The latter requirement implies that
	if two operations $o_1$ and $o_2$ from the same session are applied to two
	different replicas $r_1$ and $r_2$, the second operation cannot be discharged
	until the effect of $o_1$ is included in $r_2$. The \rcf{getBalance} operation
	requires causal consistency, as it requires the operations from the same
	session to be visible, which cannot be guaranteed under eventual consistency.
	The corresponding store semantics is captured by the contract $\ccc$ defined
	below:

@@

The quantifier in the sequent binds $\cureff$ that occurs free in $\cv_{st}$
and $\cv_{op}$. Context ($\Delta$) of the sequent is a conjunction of
assumptions about the nature of primitive relations. A \emph{well-formed}
axiomatic execution ($\E$) is expected to satisfy these assumptions (i.e., $\E
\models \Delta$).

@@

\begin{itemize}
\item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
\item Visibility only relates actions on the same object:
	\begin{itemize}
	\item $\small \forall a,b.~\vis{a}{b} \Rightarrow \sameobj{a}{b}$.
	\end{itemize}
\item Session order is a transitive relation:
	\begin{itemize}
	\item $\small \forall a,b,c.~\so{a}{b} ~\wedge~ \so{b}{c} \Rightarrow \so{a}{c}$.
	\end{itemize}
\item Same object is an equivalence relation:
	\begin{itemize}
	\item $\small \forall a.~\sameobj{a}{a}$.
	\item $\small \forall a,b.~\sameobj{a}{b} \Rightarrow \sameobj{b}{a}$.
	\item $\small \forall a,b,c.~\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} \Rightarrow \sameobj{a}{c}$.
	\end{itemize}
\end{itemize}
\end{definition}

@@

In this scheme, the consistency level of an operation is any combination of the
above guarantees, which form a partially ordered consistency lattice show in
Figure~\ref{fig:lattice}. Each element in this lattice corresponds to a
store-consistency level, and is represented by its contract. An edge from a
lower level element to an upper level element corresponds to logical
implication between their corresponding contracts. Classifying a contract under
this scheme is a directed search in the lattice, starting from the bottom, and
determining the weakest consistency level under which the contract can be
satisfied. Under this scheme, \rcf{deposit} operations does not need any
guarantees, \rcf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge
\R{WFR} \Rightarrow \cv_{gb}$), and \rcf{withdraw} cannot be satisfied ($\Delta
\vdash \R{RYW} \wedge \R{MW} \wedge \R{MR} \wedge \R{WFR} \not\Rightarrow
\cv_{w}$).

@@

We now present a meta-theoretic result that certifies the soundness of
classification-based contract enforcement. To help us state the result, we
define an operational semantics of the system described informally in
\S~\ref{sec:sysmod}:

\begin{smathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{lclcl}
{\it op} 	& \in & \mathtt{Operation} \\
{\tau}		& \in & \mathtt{Consistency Class} 	& \coloneqq & {\sf ec},{\sf cc},{\sf sc} \\
{\sigma} 	& \in & \mathtt{Session} 					 	& \coloneqq & \cdot \ALT \langle op,\tau \rangle; \sigma \\
\Sigma 		& \in & \mathtt{Session\;Soup}   	 	& \coloneqq & \sigma \pll \Sigma \ALT \emptyset \\
					&			&	\mathtt{Config}		  			 	& \coloneqq & \E,\Sigma \\
\end{array}
\end{smathpar}

We model the system as a tuple $\E,\Sigma$, where the axiomatic execution $\E$
captures the data store's current state, and session soup $\Sigma$ is the set
of concurrent client sessions interacting with the store. A session $\sigma$ is
a sequence of pairs composed of replicated data type operations $\mathit{op}$,
tagged with the consistency class $\tau$ of their contracts (as determined by
the contract classification scheme). We assume a reduction relation of form:

\begin{smathpar}
  \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E',\sigma \pll \Sigma}
\end{smathpar}

>>>>

The syntax of our core contract language is shown in Figure
~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
admits prenex universal quantification over typed and untyped effect variables.
We use a special effect variable ($\cureff$) to denote the effect of
\emph{current operation} - the operation for which a contract is being written.
Notice that $\cureff$ occurs free in the contract. We fix its scope when
classifying contracts (\S~\ref{sec:classify}). The type of an effect is
simply the name of the operation (eg: \rcf{withdraw}) that induced the effect.
We admit disjunction in types to let an effect variable range over multiple
operation names. The contract $\small \forall (a : \tau_1 \vee \tau_2).~\psi$
is just syntactic sugar for $\small \forall a. (\oper{a}{\tau_1} \vee
\oper{a}{\tau_2}) \Rightarrow \psi$. An untyped effect variable ranges over all
operation names.

Quantifier-free propositions in our contract language are conjunctions,
disjunctions and implications of predicates expressing relations between pairs
of effect variables. The syntactic class of relations is seeded with primitive
$\visZ$, $\soZ$, and $\sameobjZ$ relations, and also admits derived relations
that are expressible as union, intersection, or transitive
closure\footnote{Strictly speaking, $R^{+}$ is not the transitive closure of
$R$, as transitive closure is not expressible in FOL.  Instead, $R^{+}$ in our
language denotes \emph{a} superset of transitive closure of $R$. Formally,
$R^{+}$ is any relation $R'$ such that forall $x$, $y$, and $z$, a) $R(x,y)
\Rightarrow R'(x,y)$, and b) $R'(x,y) \conj R'(y,z) \Rightarrow R'(x,z)$.} of
primitive relations.  Commonly used derived relations are the \emph{same object
session order} ($\small \sooZ = \soZ ~\cap~ \sameobjZ$), \emph{happens-before
order} ($\small \hbZ = (\soZ ~\cup~ \visZ)^+$) and the \emph{same object
happens-before order} ($\small \hboZ = (\sooZ ~\cup~ \visZ)^+$).

@@

\item \textbf{Causal consistency}: Causally consistent operations are required
	to see a causally consistent snapshot of the object state, including the
	actions performed on the same session.  The latter requirement implies that
	if two operations $o_1$ and $o_2$ from the same session are applied to two
	different replicas $r_1$ and $r_2$, the second operation cannot be discharged
	until the effect of $o_1$ is included in $r_2$. The \rcf{getBalance} operation
	requires causal consistency, as it requires operations from the same
	session to be visible, which cannot be guaranteed under eventual consistency.
	The corresponding store semantics is captured by the contract $\ccc$ defined
	below:

@@

The quantifier in the sequent binds $\cureff$ that occurs free in $\cv_{st}$
and $\cv_{op}$. The context ($\Delta$) of the sequent is a conjunction of
assumptions about the nature of primitive relations. A \emph{well-formed}
axiomatic execution ($\E$) is expected to satisfy these assumptions (i.e., $\E
\models \Delta$).

@@

\begin{itemize}
\item The happens-before relation is acyclic: $\small \forall a.~\neg\hbo{a}{a}$.
\item Visibility only relates actions on the same object:
	\begin{itemize}
	\item $\small \forall a,b.~\vis{a}{b} \Rightarrow \sameobj{a}{b}$.
	\end{itemize}
\item Session order is a transitive relation:
	\begin{itemize}
	\item $\small \forall a,b,c.~\so{a}{b} ~\wedge~ \so{b}{c} \Rightarrow \so{a}{c}$.
	\end{itemize}
\item Same object is an equivalence relation:
	\begin{itemize}
	\item $\small \forall a.~\sameobj{a}{a}$.
	\item $\small \forall a,b.~\sameobj{a}{b} \Rightarrow \sameobj{b}{a}$.
	\item $\small \forall a,b,c.~\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} \Rightarrow \sameobj{a}{c}$.
	\end{itemize}
\end{itemize}
\end{definition}

@@

In this scheme, the consistency level of an operation is any combination of the
above guarantees, which form a partially ordered consistency lattice show in
Figure~\ref{fig:lattice}. Each element in this lattice corresponds to a
store-consistency level, and is represented by its contract. An edge from an
upper level element to a lower level element corresponds to a weaker-than
relation between the corresponding contracts. Classifying a contract under
this scheme is a directed search in the lattice, starting from the bottom, and
determining the weakest consistency level under which the contract can be
satisfied. Under this scheme, \rcf{deposit} operations do not need any
guarantees, \rcf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge
\R{WFR} \Rightarrow \cv_{gb}$), and \rcf{withdraw} cannot be satisfied ($\Delta
\vdash \R{RYW} \wedge \R{MW} \wedge \R{MR} \wedge \R{WFR} \not\Rightarrow
\cv_{w}$).

@@

We now present a meta-theoretic result that certifies the soundness of
classification-based contract enforcement. To help us state the result, we
define an operational semantics of the system described informally in
\S~\ref{sec:sysmod}:
\begin{mathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{lclcl}
{\it op} 	& \in & \mathtt{Operation} \\
{\tau}		& \in & \mathtt{Consistency Class} 	& \coloneqq & {\sf ec},{\sf cc},{\sf sc} \\
{\sigma} 	& \in & \mathtt{Session} 					 	& \coloneqq & \cdot \ALT \langle op,\tau \rangle; \sigma \\
\Sigma 		& \in & \mathtt{Session\;Soup}   	 	& \coloneqq & \sigma \pll \Sigma \ALT \emptyset \\
					&			&	\mathtt{Config}		  			 	& \coloneqq & \E,\Sigma \\
\end{array}
\end{mathpar}

We model the system as a tuple $\E,\Sigma$, where the axiomatic execution $\E$
captures the data store's current state, and session soup $\Sigma$ is the set
of concurrent client sessions interacting with the store. A session $\sigma$ is
a sequence of pairs composed of replicated data type operations $\mathit{op}$,
tagged with the consistency class $\tau$ of their contracts (as determined by
the contract classification scheme). We assume a reduction relation of form:
\begin{mathpar}
  \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E',\sigma \pll \Sigma}
\end{mathpar}

### 2015-04-10 evaluation.tex
edits

The distribution of contracts in these applications is given in
Table~\ref{tab:ctrts}. We see that majority of the operations and transactions
are classified as eventually consistent and RC, respectively. Operation
contracts are used to enforce integrity and visibility constraints on
individual fields in the tables. Transactions are mainly used to consistently
modify and access related fields across tables. In \name, the contract
classification process is completely performed at compile time and has no
overheads at runtime. The proof obligations associated with contract
classification is discharged through the Z3 SMT Solver. Across our benchmarks,
classifying a contract took 11.5 milliseconds on average.

For our performance evaluation, we deploy \name applications in
\emph{clusters}, where each cluster is composed of five fully replicated
Cassandra replicas within the same datacenter. We instantiate one shim layer
node co-located on the same VM as a Cassandra replica. Clients are instantiated
within the same data center as the store, and run transactions. We deploy each
cluster and client node on an \cf{c3.4xlarge} Amazon EC2 instance. We call this
a \cf{1DC} configuration. For our geo-distributed experiments (\cf{2DC}), we
instantiate 2 clusters, each with five nodes, and place the clusters on US-east
(Virginia) and US-west (Oregon) locations. The average inter-region latency was
85ms.

>>>>

The distribution of contracts in these applications is given in
Table~\ref{tab:ctrts}. We see that majority of the operations and transactions
are classified as eventually consistent and RC, respectively. Operation
contracts are used to enforce integrity and visibility constraints on
individual fields in the tables. Transactions are mainly used to consistently
modify and access related fields across tables. In \name, the contract
classification process is completely performed at compile time and incurs no
runtime overhead. The proof obligations associated with contract
classification is discharged through the Z3 SMT Solver. Across our benchmarks,
classifying a contract took 11.5 milliseconds on average.

For our performance evaluation, we deploy \name applications in
\emph{clusters}, where each cluster is composed of five fully replicated
Cassandra replicas within the same datacenter. We instantiate one shim layer
node co-located on the same VM as a Cassandra replica. Clients are instantiated
within the same data center as the store, and run transactions. We deploy each
cluster and client node on a \cf{c3.4xlarge} Amazon EC2 instance. We call this
a \cf{1DC} configuration. For our geo-distributed experiments (\cf{2DC}), we
instantiate 2 clusters, each with five nodes, and place the clusters on US-east
(Virginia) and US-west (Oregon) locations. The average inter-region latency was
85ms.

### 2015-04-10 implementation.tex
Fixing figures.
edits

\begin{figure}
\begin{center}
\includegraphics[width=0.7\columnwidth]{Figures/AtomicityImpl}
\end{center}
\caption{Implementing atomicity semantics. Dotted circle represents effects not
yet inserted into the backing store.}
\label{fig:atomicity_impl}
\end{figure}

\name implements atomic visibility by exploiting shim layer causality guarantee
-- an effect is included only if all the effects if depends on are also
included. Consider the example given in Figure~\ref{fig:atomicity_impl}. For
every transaction in \name, we instantiate a special transaction marker
effect $m$. But importantly, do not insert into the backing store. $m$ is
included as a dependence to every effect generated in the transaction. In the
figure, the graph on the left shows the state of the store in the middle of a
transaction. Each circle represents an effect. The dotted circle indicates that
the effect has been instantiated, but has not yet been inserted into the store.
Since the causally preceding effect $m$ has not yet been written to the store,
no operation will witness $e1$ and $e2$ while the transaction in progress.
After the transaction has finished execution, we insert $m$ into the backing
store, marking all the effects from the transactions as a dependence for $m$.
Now any replica which includes one of the effects from the transaction must
include $m$, and transitively must include every effect from the transaction.
This ensures atomicity and satisfies the RC requirement.

The above scheme prevents a transaction from witnessing its own effects. This
might conflict with the causality requirement on the operations. Hence,
transactions piggy-back the previous effects from the same transaction for each
request. MAV semantics is implemented by keeping track of the set of
transaction markers $M$ witnessed by the transaction, and before performing an
operation at some replica, ensuring that $M$ is a subset of the transaction
markers included at that replica. If not, the missing effects are synchronously
fetched. RR semantics is realized by capturing a optimized snapshot of the
state of some replica; each operation from an RR transaction is applied to this
snapshot state. Any generated effects are added to this snapshot.

>>>>

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{Figures/AtomicityImpl}
\end{center}
\caption{Implementing atomicity semantics. Dotted circle represents effects not
yet inserted into the backing store.}
\label{fig:atomicity_impl}
\end{figure}

\name implements atomic visibility by exploiting shim layer causality guarantees
-- an effect is included only if all the effects if depends on are also
included. Consider the example given in Figure~\ref{fig:atomicity_impl}. For
every transaction in \name, we instantiate a special transaction marker
effect $m$ that is importantly not inserted into the backing store. Marker $m$ is
included as a dependence to every effect generated in the transaction. In the
figure, the top graph on the left shows the state of the store in the middle of a
transaction. Each circle represents an effect. The dotted circle indicates that
the effect has been instantiated, but has not yet been inserted into the store.
Since the causally preceding effect $m$ has not yet been written to the store,
no operation will witness $e1$ and $e2$ while the transaction in progress.
After the transaction has finished execution, we insert $m$ into the backing
store, marking all the effects from the transactions as a dependence for $m$.
Now, any replica which includes one of the effects from the transaction must
include $m$, and transitively must include every effect from the transaction.
This ensures atomicity and satisfies the RC requirement.

The above scheme prevents a transaction from witnessing its own effects. This
might conflict with causality requirements on the operations. Hence,
transactions piggy-back the previous effects from the same transaction for each
request. MAV semantics is implemented by keeping track of the set of
transaction markers $M$ witnessed by the transaction, and before performing an
operation at some replica, ensuring that $M$ is a subset of the transaction
markers included at that replica. If not, the missing effects are synchronously
fetched. RR semantics is realized by capturing a optimized snapshot of the
state of some replica; each operation from an RR transaction is applied to this
snapshot state. Any generated effects are added to this snapshot.

### 2015-04-10 intro.tex
edits

Many real-world web services --- such as those built and maintained by Amazon,
Facebook, Google, Twitter, etc. --- replicate application state and logic
across multiple \emph{replicas} within and across data centers. Replication is
intended not only to improve application throughput and reduce user-perceived
latency, but also to tolerate partial failures without compromising overall
service availability. Traditionally programmers have relied on \emph{strong
consistency} guarantees such as linearizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct applications.
While strong consistency is an easily stated property, it masks the reality
underlying large-scale distributed systems with respect to non-uniform latency,
availability and network partitions~\cite{Brewer2000,Gilbert2002}. Indeed,
modern web services, which aim to provide an "always-on" experience,
overwhelmingly favor availability and partition tolerance over strong
consistency. To this end, several \emph{weak consistency} models such as
eventual consistency, causal consistency, session guarantees, and timeline
consistency have been proposed.

@@

\begin{itemize}[noitemsep]
\item We introduce \name, a shallow extension of Haskell that supports the
	description and validation of replicated data types found on \ecds. Contracts
	are used to specify fine-grained application-level consistency properties,
	and are statically analyzed to assign the most efficient and sound store
	consistency level to the corresponding operation.
\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct isolation level for a transaction.
\item We provide meta-theory that certifies the soundness of our contract
	enforcement system, and ensures that an operation is only executed if the
	required conditions on consistency are met.
\item An implementation of \name as a transparent shim layer over
	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.
	Experimental evaluation over a set of real-world applications, including a
	Twitter-like micro-blogging site and an eBay-like auction site illustrates
	the practicality of our approach.
\end{itemize}

\noindent The rest of the paper is organized as follows. The next section
describes the system model.  We describe the challenges in programming under
eventual consistency, and introduce \name\ contracts as a proposed solution to
overcome these issues in \S~\ref{sec:motivation}. \S~\ref{sec:lang} provides
more details on the contract language, and its mapping to the store consistency
levels, along with meta-theory for certifying the correctness of the mapping.
\S~\ref{sec:txns} introduces transaction contracts and their classification.
\S~\ref{sec:impl} describes the implementation of \name on top of Cassandra.
\S~\ref{sec:results} discusses experimental evaluation. \S~\ref{sec:related}
and~\ref{sec:concl} present related work and conclusions.

>>>>

Many real-world web services --- such as those built and maintained by Amazon,
Facebook, Google, Twitter, etc. --- replicate application state and logic
across multiple \emph{replicas} within and across data centers. Replication is
intended not only to improve application throughput and reduce user-perceived
latency, but also to tolerate partial failures without compromising overall
service availability. Traditionally programmers have relied on \emph{strong
consistency} guarantees such as linearizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct applications.
While strong consistency is an easily stated property, it masks the reality
underlying large-scale distributed systems with respect to non-uniform latency,
availability, and network partitions~\cite{Brewer2000,Gilbert2002}. Indeed,
modern web services, which aim to provide an "always-on" experience,
overwhelmingly favor availability and partition tolerance over strong
consistency. To this end, several \emph{weak consistency} models such as
eventual consistency, causal consistency, session guarantees, and timeline
consistency have been proposed.

@@

\begin{itemize}
\item We introduce \name, a shallow extension of Haskell that supports the
  description and validation of replicated data types found in an
  \ecds. Contracts are used to specify fine-grained application-level
  consistency properties, and are statically analyzed to assign the most
  efficient and sound store consistency level to the corresponding
  operation.
\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct isolation level for a transaction.
\item We provide meta-theory that certifies the soundness of our contract
	enforcement system, and ensures that an operation is only executed if the
	required conditions on consistency are met.
\item We describe an implementation of \name as a transparent shim layer over
	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.
	Experimental evaluation over a set of real-world applications, including a
	Twitter-like micro-blogging site and an eBay-like auction site illustrates
	the practicality of our approach.
\end{itemize}

\noindent The rest of the paper is organized as follows. The next section
describes the system model.  We describe the challenges in programming under
eventual consistency, and introduce \name\ contracts as a proposed solution to
overcome these issues in \S~\ref{sec:motivation}. \S~\ref{sec:lang} provides
more details on the contract language, and its mapping to store consistency
levels, along with meta-theory for certifying the correctness of the mapping.
\S~\ref{sec:txns} introduces transaction contracts and their classification.
\S~\ref{sec:impl} describes the implementation of \name on top of Cassandra.
\S~\ref{sec:results} discusses experimental evaluation. \S~\ref{sec:related}
and~\ref{sec:concl} present related work and conclusions.

### 2015-04-10 motivation.tex
edits

A key novelty in \name is that it allows the addition of new RDTs to the store,
which obviates the need for coercing the application logic to utilize the store
provided data types. In addition, \name treats the convergence semantics (i.e.,
\emph{how} conflicting updates are resolved) of the data type separately from
its consistency properties (i.e., \emph{when} updates become visible). This
separation of concerns permits \emph{operational} reasoning for conflict
resolution, and \emph{declarative} reasoning for consistency. The combination
of these techniques enhances the programmability of the store.

@@

This notion of observable equivalence can be generalized to other RDTs as well.
For example, a last-writer-wins register with multiple updates is equivalent to
a register with only the last write. Similarly, a set with a collection of add
and remove operations is equivalent to a set with a series of additions of live
elements from the original set. Since the notion of observable equivalence is
specific to each RDT, programmers can provide a summarization
function - (\cf{summarize}) of type \cf{[e] -> [e]} -  as a part of
the RDT specification. The summarization function for the bank account is:

@@

Consider the execution shown in Figure~\ref{fig:unsafeWithdrawAnomaly}. Assume
that all operations in the figure are on the same bank account object with the
initial balance being zero. Session 1 performs a \rcf{deposit} of 100, followed
by a \rcf{withdraw} of 80 in the same session. The \rcf{withdraw} operation
witnesses the deposit and succeeds\footnote{Although visibility and session
order relations relate effects, we have abused the notation in these examples
to relate operations, with the idea that the relations relate the effect
emitted by those operations}. Subsequently, session 2 perform a \rcf{withdraw}
operation, but importantly, due to eventual consistency, only witnesses the
\rcf{deposit} from session 1, but not the subsequent withdraw. Hence, this
\rcf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
constraint. A subsequent \rcf{getBalance} operation, that happens to witness all
the previous operations, would report a negative balance.

@@

Although it is easy to understand the reasons behind the occurrence of the
aforementioned anomalies, finding the appropriate fixes is not readily
apparent. Making \rcf{getBalance} a strongly consistent operation is definitely
sufficient to avert anomalies, but is it really necessary? Given the cost of
enforcing strong consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid
it unless there are no viable alternatives. Exploring the space of these
alternatives requires understanding the subtle differences in semantics of
various kinds of weak consistency alternatives.

@@

In our running example, it is clear that in order to preserve the integrity
constraint, the \rcf{withdraw} operation must be strongly consistent.  That is,
given two \rcf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
vice-versa. We express this application-level consistency requirement as a
contract ($\cv_w$) over \rcf{withdraw}:

@@

To prevent \rcf{getBalance} from ever showing a negative balance, it is
necessary to prevent the scenario depicted in
fig.~\ref{fig:negativeBalanceAnomaly}. Let $\cureff$ stand for the effect
emitted by the \rcf{getBalance} operation. If the effect ($b$) of a withdraw
operation is visible to $\cureff$, and the effect ($a$) of a deposit operation
is visible to the effect ($b$) of the withdraw operation, then it must be the
case that $a$ is also visible to $\cureff$. A contract ($\cv_{gb}^1$) for
\rcf{getBalance} operation that precisely captures this application-level consistency
requirement can be written thus:

@@

\noindent To prevent the missing update anomaly described in
fig.~\ref{fig:missingUpdateAnomaly}, it is necessary for a \rcf{getBalance}
operation on a bank account to witness the effects of all previous \rcf{deposit}
and \rcf{withdraw} operations performed on the same bank account in the same
session. We can express an additional contract ($\cv_{gb}^2$) for
\rcf{getBalance} that captures this consistency requirement:

@@

\noindent Intuitively, this prevents both the \rcf{getBalance} anomalies
described in figs.~\ref{fig:negativeBalanceAnomaly} and
\ref{fig:missingUpdateAnomaly} from ever occurring.

>>>>

A key novelty in \name is that it allows the addition of new RDTs to the
store, which obviates the need for coercing application logic to utilize
store-provided data types. In addition, \name treats the convergence
semantics (i.e., \emph{how} conflicting updates are resolved) of the data
type separately from its consistency properties (i.e., \emph{when} updates
become visible). This separation of concerns permits \emph{operational}
reasoning for conflict resolution, and \emph{declarative} reasoning for
consistency. The combination of these techniques enhances the
programmability of the store.

@@

This notion of observable equivalence can be generalized to other RDTs as well.
For example, a last-writer-wins register with multiple updates is equivalent to
a register with only the last write. Similarly, a set with a collection of add
and remove operations is equivalent to a set with a series of additions of live
elements from the original set. Since the notion of observable equivalence is
specific to each RDT, programmers can provide a summarization
function - of type \cf{[e] -> [e]} -  as a part of
the RDT specification. The summarization function for the bank account is:

@@

Consider the execution shown in Figure~\ref{fig:unsafeWithdrawAnomaly}. Assume
that all operations in the figure are on the same bank account object with the
initial balance being zero. Session 1 performs a \rcf{deposit} of 100, followed
by a \rcf{withdraw} of 80 in the same session. The \rcf{withdraw} operation
witnesses the deposit and succeeds\footnote{Although visibility and session
order relations relate effects, we have abused the notation in these examples
to relate operations, with the idea that the relations relate the effect
emitted by those operations.}. Subsequently, session 2 perform a \rcf{withdraw}
operation, but importantly, due to eventual consistency, only witnesses the
\rcf{deposit} from session 1, but not the subsequent withdraw. Hence, this
\rcf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
constraint. A subsequent \rcf{getBalance} operation, that happens to witness all
the previous operations, would report a negative balance.

@@

Although it is easy to understand the reasons behind the occurrence of the
aforementioned anomalies, finding the appropriate fixes is not readily
apparent. Making \rcf{getBalance} a strongly consistent operation is
definitely sufficient to avert anomalies, but is it really necessary? Given
the cost of enforcing strong consistency~\cite{DynamoDB, Pileus}, it is
preferable to avoid imposing such stringent conditions unless there are no
viable alternatives. Exploring the space of these alternatives requires
understanding the subtle differences in semantics of various kinds of weak
consistency alternatives.

@@

In our running example, it is clear that in order to preserve the critical
integrity constraint, the \rcf{withdraw} operation must be strongly
consistent.  That is, given two \rcf{withdraw} operations $a$ and $b$,
either $a$ is visible to $b$ or vice-versa. We express this
application-level consistency requirement as a contract ($\cv_w$) over
\rcf{withdraw}:

@@

To prevent \rcf{getBalance} from ever showing a negative balance, it is
necessary to prevent the scenario depicted in
Figure~\ref{fig:negativeBalanceAnomaly}. Let $\cureff$ stand for the effect
emitted by the \rcf{getBalance} operation. If the effect ($b$) of a withdraw
operation is visible to $\cureff$, and the effect ($a$) of a deposit operation
is visible to the effect ($b$) of the withdraw operation, then it must be the
case that $a$ is also visible to $\cureff$. A contract ($\cv_{gb}^1$) for
\rcf{getBalance}  that precisely captures this application-level consistency
requirement can be written thus:

@@

\noindent To prevent the missing update anomaly described in
Figure~\ref{fig:missingUpdateAnomaly}, it is necessary for a \rcf{getBalance}
operation on a bank account to witness the effects of all previous \rcf{deposit}
and \rcf{withdraw} operations performed on the same bank account in the same
session. We can express an additional contract ($\cv_{gb}^2$) for
\rcf{getBalance} that captures this consistency requirement:

@@

\noindent Intuitively, this specification prohibits both the
\rcf{getBalance} anomalies described in
Figures.~\ref{fig:negativeBalanceAnomaly} and \ref{fig:missingUpdateAnomaly}
from  occurring.
